XXXXX

### 2014-12-15 paper/paper 5/paper 5.tex
initial commit

\newenvironment{gcotable-1}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context concept} & \textbf{p. list} & \textbf{concepts} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

\newenvironment{gcotable}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l|l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context concept} & \textbf{left p. list} & \textbf{right p. list} & \textbf{concepts} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

\newenvironment{DL}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l}
	\textbf{DL:} \\

@@

\section{Introduction}

@@

At-least restrictions can be expressed either generically by description logics (DL) or specifically by domain-specific constraint languages (DSCL) such as OWL 2, DSP, and ShEx.
With at-least restrictions we can state that having at least 2 female childs is equivalent to having at least 2 daughters. This axiom may be represented by DL in a logical way:

\begin{align*}
Having2Daughters \equiv \geq 2 hasChild.Woman
\end{align*}

The same axiom may also be represented by different DSCLs like OWL 2, DSP, and ShEx:

\begin{ex}
# OWL 2
# -----
:Having2Daughters
    a owl:Restriction ;
    owl:minQualifiedCardinality "2"^^xsd:nonNegativeInteger ;
    owl:onProperty :hasChild ;
    owl:onClass :Woman .

# DSP
# ---			
:descriptionTemplate 
    a dsp:DescriptionTemplate ; 
    dsp:resourceClass :Having2Daughters ; 
    dsp:statementTemplate [
        a dsp:NonLiteralStatementTemplate ;
        dsp:minOccur "1"^^xsd:nonNegativeInteger ; 
        dsp:maxOccur "infinity"^^xsd:string ; 
        dsp:property :hasChild ; 
        dsp:nonLiteralConstraint [ 
            a dsp:NonLiteralConstraint ;
            dsp:valueClass :Woman ] ] .

# ShEx
# ----
<Having2Daughters> {         
    :hasChild @<Woman>{2, }
}
<Woman> {         
    rdf:type (:Woman)
}
\end{ex}

Data, which is valid according to the specified at-least restriction, looks like this:

\begin{ex}
:Marge
    a :Having2Daughters , :Woman ;
    :hasChild :Maggie , :Lisa .
:Maggie
    a :Woman .
:Lisa
    a :Woman .
\end{ex}

If :Marge has only 1 child and if :Marge is assigned to the class :Having2Daughters, a constraint violation is raised for the following invalid data:

\begin{ex}
:Marge
    a :Having2Daughters , :Woman ;
    :hasChild :Lisa .
:Lisa
    a :Woman .
\end{ex}

As there is no standard constraint language, constraint designers may choose different DSCLs in order to express the same at-least restriction. 
When constraint designers choose one DSCL in order to express an arbitrary constraint, it should be possible to transform this constraint into a constraint expressed by any other DSCL. 
This is important as it is often the case that one constraint designer knows to read and to write constraints in one DSCL very well but not in another DSCL. 
This way, constraint designers may communicate with each other without the necessity to understand the other DSCL.

There are validation environments available enabling to automatically validate any RDF constraints expressed by just a few DSCL such as OWL 2 and DSP.
But it should be possible to validate any RDF constraints expressed by any DSCL automatically and what is even more important in the same manner without any differences.

The following table contains the mapping between the at-least restriction expressed by description logics and the at-least restriction expressed by a generic constraint language conforming to an ontology describing constraint languages.
Each constraint is either a constraint on classes or a constraint on properties.

\begin{gcotable}
property & Having2Daughters & hasChild & - & Woman & $\geq$ & 2 \\
\end{gcotable}

The at-least restriction (expressed by the generic constraint language) is represented in RDF as follows:

\begin{ex}
:minCardinalityConstraint
    a gclo:PropertyConstraint ;
    gclo:contextConcept :Having2Daughters ;
    gclo:propertyList ( :hasChild ) ;
    gclo:concept :Woman ;
    gclo:constrainingElement "\geq" ;
    gclo:constrainingValue "2" .
\end{ex}

We implement an automatic validation for any constraint expressed by a generic constraint language conforming to an ontology describing constraint languages.
If we transform a specific constraint (expressed by any DSCL) into a generic constraint (expressed by the generic constraint language), we will be able to validate the generic constraint and therefore the specific constraint immediately. 
As a consequence, we do not have to specify an automatic validation for each specific constraint, as we need to define an automatic validation once for each generic constraint. The underlying SPARQL query is executed within the SPIN validation environment:

\begin{ex}
# constraint
# ----------
?propertyConstraint
    a gclo:PropertyConstraint ;
    gclo:contextConcept ?contextConcept ;
    gclo:propertyList ( ?property ) ;
    gclo:concept ?concept ;
    gclo:constrainingElement ?constrainingElement ;
    gclo:constrainingValue ?constrainingValue .

# data
# ----
?this ?property ?o .
?this rdf:type ?contextConcept .

# validation ( at-least restriction )
# -----
FILTER ( ?constrainingElement = "\geq" ) .
BIND ( owl2spin:qualifiedCardinality( ?this, ?property, ?concept ) AS ?c ) .
BIND( STRDT ( STR ( ?constrainingValue ), xsd:nonNegativeInteger ) AS ?MinimumCardinality )
BIND( STRDT ( STR ( ?c ), xsd:nonNegativeInteger ) AS ?cardinality )
FILTER ( ?cardinality < ?MinimumCardinality ) .		  
\end{ex}

When domain experts use graphical user interfaces to define constraints in a user-friendly way, 
these constraints can be either mapped to either a specific or a generic constraint, as for both an automatic validation is provided in the background.    
So, the user does not have to know how to express specific constraints using a DSCL such as OWL 2, DSP, or ShEx.
If there is a new constraint not covered by any DSCL you can just define the automatic validation for this newly defined constraint.
This way, you are able to extend existing DSCL and to define new DSCL if existing DSCLs are not sufficient for further use cases.

Our \textbf{contributions} of this paper are:
\begin{enumerate}
	\item each constraint (expressed by any DSCL) can be validated automatically
	\item any specific constraint (expressed by any DSCL) can be transformed into a generic constraint (expressed by a generic constraint language)
	\item we developed an ontology describing any constraint language
\end{enumerate}

Benefits:
\begin{itemize}
	\item GCL with only 6 elements
	\item if a language designer develops a new DSCL the language designer only needs to define a mapping between the new DSCL and the GCLO 
	in order to be able to automatically validate the new DSCL constraints
\end{itemize}

idea:
\begin{itemize}
	\item each specific constraint expressed using any domain specific language can be expressed as DL constraint using DL
	\item each DL constraint can be expressed as generic constraint using generic constraint language
	\item meta-model based ontology describing constraints
	\item possible limitation: maybe only typical constraints
\end{itemize}

purpose:
\begin{itemize}
	\item translations of constraints expressed using any domain specific constraint language
\end{itemize}

contributions:
\begin{itemize}
	\item 
\end{itemize}

assumption:
\begin{itemize}
	\item CWA
\end{itemize}

ontology design pattern.org
constraint grammar
constraint elements are the same
DSP uses other design pattern / 
constraint / constraint elements / constraint design patterns / constraint language
define terminology
related to requirements DB / relate requirements to constraints, constraint languages
constraint element: e.g. min cardinality
dependency between requirements / when this requirements is fulfilled then this is also fulfilled (e.g. min card and requ. car)
min car more powerful than req. propery
why ontology? automatic translation from DSP to OWL 
define general constraint language
we do not invent a new constraint languages
requirements in SPARQL umsetzten
only one validation for each requirement
constraints may be composed

\begin{itemize}
	\item any specific constraint (expressed by DSCL) must be mapped to a generic constraint (expressed by GCL)
\end{itemize}

\section{An Ontology Describing Constraint Languages}

fig:generic constraint language ontology.  
\begin{figure}
	\centering
		\includegraphics[width=1.00\textwidth]{generic constraint language ontology/generic constraint language ontology.png}
	\caption{generic constraint language ontology}
	\label{fig:generic constraint language ontology}
\end{figure}

terminology:
\begin{itemize}
	\item A "constraint" is an actual constraint, like dcterms:creator :hasMinCardinality "1" .
  \item a "constraint language" like DSP is used to formulate constraints.
  \item a constraint language consists of "constraint elements" like :hasMinCardinality (this is probably the most controverse definition).
	\item class constraint: constraint on context class
\end{itemize}

\begin{itemize}
	\item relationship to requirement
  \item ontology describing constraint languages
	\item interoperability between constraint languages
  \item express constraint in CL A and translate constraint in CL B
  \item provide SPARQL validation (SPIN mapping) for each general C 
\end{itemize}

>>>>

\newenvironment{gcotable-1}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}\linebreak
  \begin{tabular}{l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context concept} & \textbf{p. list} & \textbf{concepts} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

\newenvironment{gcotable}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}\linebreak
  \begin{tabular}{l|l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context concept} & \textbf{left p. list} & \textbf{right p. list} & \textbf{concepts} & \textbf{c. element} & \textbf{c. value} \\
  \hline

@@

\newenvironment{DL}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}\linebreak
  \begin{tabular}{l}
	\textbf{DL:} \\

@@

\begin{itemize}
	\item check quality of thesauri using validation
\end{itemize}

\textbf{validations:}
\begin{itemize}
	\item how far the thesaurus conforms to established ISO norms for thesauri
	\item hierarchical relations between terms and relations
	\item classification of terms
	\item check links to other thesauri: STW Thesaurus for Economics of the ZBW, AGROVOC thesaurus of FAO, DBpedia
	\item typical knowledge modeling patterns occurring by the use of SKOS and SKOS-XL
	\item ambiguous terms: equivalence and compound equivalence relationships between preferred and non-preferred terms
	\item representation of compound relationships and compound concept/ compound concepts are part of the ISO2788 standard
\end{itemize}

\textbf{validation of datasets conversion to SKOS:}
\begin{itemize}
	\item are terms and relations mapped adequately to SKOS class and properties? 
	\item completeness
	\item relations between generic and ambiguous terms (having different meanings in specialized contexts) 
\end{itemize}

\textbf{TheSoz specific validations:}
\begin{itemize}
	\item TheSoz extensions / own namespace / TheSoz specific classes and properties
\end{itemize}

\section{Validations}

\textbf{map term to multiple terms not possible in SKOS:}
A concept in one thesaurus
might correspond to a combination of two concepts in
another thesaurus, e.g. the term "Electronic Government" of the TheSoz has originally been mapped to
the combination of the terms "Public Administration"
and "Internet" of the STW. The mapping properties of
SKOS do not allow such single-to-multiple relations
(neither for one language nor for multiple languages).

@@

\section{Visualization Ideas}

ich erstelle ja constraint violation triples, die so aussehen:

\begin{ex}
CONSTRUCT {
    _:violation
        a spin:ConstraintViolation ;
        rdfs:label ?violationMessage ;
        spin:violationRoot ?violationRoot ;
        spin:violationPath ?violationPath ;
        spin:violationSource ?violationSource ;
        spin:fix ?violationFix ;
        :validationLevel :error }
\end{ex}

Auch den validation level möchte ich angeben, da wird es sicher was geben, zumindest irgendwann.
Ich nutzte in der zwischenzeit spin:validationLevel obwohl es das im spin namespace (noch) nicht gibt, würde aber sehr viel soinn machen.

Wir haben dann als ergebnis der validierung einen graphen mit constraint validation tripeln.
diese kann man nach validation level (e.g. error, fail, message, ...) sortieren, unterteilen und auch getrennt visualisieren.

use case:
visualisierung nur der wirklich wirklich wichtigen (fail) constraint violation triples

use case:
visualisierung einer bestimmten constraint violation triple gruppe

-----
da ich immer auch den root, also das subject von constraint violations mit generiere, kann man auch eine visualisierung erstellen bestimmter rdf subjects, die der grund für constraint violations darstellen, wieder sortiert nach validation level

----
soertierung nach violation source ist auch denkbar, also welche constraint ausgedrückt durch welche DSCL hat welche (wieviele) constraint violations verursacht (spin:violationSource).

----

spin:fix zeigt auf triple die angeben, wie eine constraint violation behoben werden kann, auch das kann visualisiert werden.
z.b. eine kette von aktionen, die dazu führen, dass bestimmte constraint violations behoben werden können.

--

das sprudeln die ideen...

-----

spin:path --> welche properties sind an einem constraint violation beteiligt. hier könnte man einen pfad aufzeigen, der für eine bestimmte constraint violation zeigt, welche properties in welcher reihenflge beteiligt sind. eine property chain

--

 also mit all diesen generierten informationen lässt sich viel anfangen und sinnvolle dinge visualisieren.
wir können auch mal drüber sprechen.

---
macht es sinn auch die constraints selber zu visualisieren?

- welche constraints gibt es die für eine bestimmte DSCL in SPIN definiert wurden?
- wenn eine specific constraint aug eine generic constraint gemappt wird, können wir visualisieren wie spezifische constraints zueinander in beziehung stehen und auch zu anderen spezifischen constraints anderer DSCLs
- macht eine visualisierung der SPARQL queries sinn die verwendet werden constraints zu validieren?
- welche specific und generic constraints sind welchen requirements zugeordnet, kömnte auch visualisiert werden...

ich denke das reicht jetzt erst mal....
